Compositional equivalences based on Open pNets

### 2020-05-12 2016-Weak-FH-Bisim/WeakBisim.tex
appendix

\Post\,'=\Post\uplus\bigotimes_{p=n}^{0}
		\Post_p
\end{mathpar}
Then, we have
	\[ P[Q]_{j_0}  
	\models
	{\openrule
		{
			\gamma_j^{j\in (J\setminus\{j_0\}) \uplus J_Q}, 
			\Pred\,',  \Post\,'
			 }
		{\ostate{s_i^{i\in L\uplus L_Q}} \OTWeakarrow {\alpha}
			\ostate{s_i'^{\, i\in L\uplus L_Q}}}
	}
	\]
\end{lemma}
~~\\
\begin{proof} Suppose we have:
\begin{mathpar}
%\[
P\models{\openrule
	{
		\gamma_j^{j\in J}, 
		\Pred,  
		\Post}
	{\ostate{s_i^{i\in L}} \OTWeakarrow {\alpha}
		\ostate{s_i'^{\, i\in L}}}
}
\end{mathpar}
By Lemma \ref{lem-rel-OT-WOT} this implies the following: \\			
%\begin{mathpar}
\begin{small}
$\forall p\!\in\![0..m_1].P\models\openrule
    {{\beta_{1pj}^{j\in J_{1p}}},\Pred_{1p},\Post_{1p}   }
         {\ostate{{(s_{1pi})}^{i\in L}} \OTarrow {\tau} \ostate{{(s_{1(p+1)i})}^{i\in L}}},  \quad
P\models\openrule
         {\beta_{2j}^{j\in J_2},\Pred_2,\Post_2 }
         {\ostate{({s_{20i}})^{i\in L}} \OTarrow {\alpha'} \ostate{({s_{21i}})^{i\in L}}} $ 
\\
and  $\forall p \in [0..m_2].P\models\openrule
         {
           \beta_{3pj}^{j\in J_{3p}},\Pred_{3p}, \Post_{3p}}
         { \ostate{(s_{3pi})^{i\in L}} \OTarrow {\tau} \ostate{(s_{3(p+1)i})^{i\in L}}}$
\end{small}
%\end{mathpar}

@@

However it must be noticed that in steps 1 and 3, we have two kinds of WOTs with different signatures (depending on whether $Q$ moves or not). It is still possible to glue them together in a global rules with two more terms for $\Pred$ and $\Post$ terms. This global merge is possible because the post-conditions of $P$ only act on variables of $P$ and those of $Q$ on variables of $Q$ (for example $\Post_i$ has no effect on $\Pred_{1p}$ and thus does not need to be taken into account when dealing with WOTs where $Q$ does not move).

We now  compare each element of the obtained WOT with the conclusion of the lemma:
\begin{align*}
\alpha''&=\alpha'\displaystyle{\subst{\bigotimes_{j=m_1}^{0}\Post\,'_{1j}}}\\
& = \alpha'\displaystyle{\subst{\bigotimes_{j=m_1}^{0}\Post_{1j}}}&\text{$\alpha'$ only contains variables of $P$ untouched by $\Post_i$}\\
& = \alpha
\end{align*}

For $\set{\gamma'}$ we distinguish elements in the holes of $P$ and of $Q$. First suppose $j\in J\setminus\{j_0\}$ we have $\gamma'_j=\gamma_j$ because $\Post'_{ij}$ has no effect on variables of $P$ and on $\beta_{1pj}$, consequently\\
\begin{equation*}
\begin{split}\gamma'_j=&\mybigdotcup_{i=0}^{m_1} \vis{{\beta_{1ij}}\subst{\bigotimes_{k=i-1}^{0} \Post_{1k} } }  \dotcup  \vis{{\beta_{2j}}\subst{\bigotimes_{k=m_1}^{0}\Post_{1k}}}\, \dotcup\\
&
\mybigdotcup_{i=0}^{m_2} \vis{\beta_{3ij} \subst{\bigotimes_{k=i-1}^{0}\Post_{3k}\shortotimes\Post_2\shortotimes\bigotimes_{k=m_1}^{0}\Post_{1k}}}
\end{split}
\end{equation*}

Now, when $j\in J_t$ for some $t$,  $\gamma'_j$ is the concatenation of elements of $\gamma'_{1ij}$, $\gamma'_{2 j}$, $\gamma'_{3ij}$  that are not empty. By construction the concatenation of these elements is $\gamma_{tj}$, for $t\in[0..n]$. $\Post_{ik}$ has no effect on $\gamma_{tj}$ but  $\Post_{k}$ has.
%\TODO{do we want to be more explicit in the explanation? giving the exact indices is cumbersome because Post'ij is either empty or postij for the same indices as gamma, this can also be related with the ni and n'i}
We obtain:
\begin{equation*}
\begin{split}
\gamma'_j =&
\mybigdotcup_{i=0}^{m_1}{\gamma'_{1ij}}\subst{\bigotimes_{k=i-1}^{0}\Post\,'_{1k}}\dotcup
 {\gamma'_{2 j}}\subst{\bigotimes_{k=m_1}^{0}\Post\,'_{1k}}\,\dotcup\\
& \mybigdotcup_{i=0}^{m_2}{\gamma'_{3ij}}\subst{\bigotimes_{k=i-1}^{0}\Post\,'_{3k}\shortotimes\Post\,'_2\shortotimes\bigotimes_{k=n}^{0}\Post\,'_{1k}}~\dotcup~
{\mybigdotcup_{t=0}^{n} {\gamma_{tj}}\subst{\bigotimes_{k=t-1}^{0} \Post_{k}}}   
\end{split}
\end{equation*}

Concerning predicates, we also separate predicates on $P$ from predicates on $Q$, and from the equality on the action filling the hole:\\
%\begin{small}
%\begin{equation*}
%\begin{split}
%\Pred\,''&=\bigwedge_{i=0}^{m_1}\Pred\,'_{1i}\subst{\bigotimes_{j=i-1}^{0}\Post\,'_{1j}}\land\Pred\,'_2 \subst{\bigotimes_{j=m_1}^{0}\Post\,'_{1j}}\land\\
%&~~{\bigwedge_{i=0}^{m_3}\Pred\,'_{3i}\subst{\bigotimes_{j=i-1}^{0}\Post\,'_{3j}\shortotimes\Post\,'_2\shortotimes\bigotimes_{j=m_1}^{0}\Post\,'_{1j}}}\\
%&=\big(\bigwedge_{p=0}^{m_1}\!\Pred_{1p}\subst{\bigotimes_{j=p-1}^{0}\Post_{1j}}\land\Pred_2\subst{\bigotimes_{p=m_1}^{0}\Post_{1p}}\land\\
%&~~~\bigwedge_{p=0}^{m_2}\!\Pred_{3p}\subst{\bigotimes_{j=p-1}^{0}\Post_{3j}\otimes\Post_2 \otimes\bigotimes_{p=m_1}^{0}\Post_{1p}}\big) \land \bigwedge_{t=0}^{n} \Pred_t\bigotimes_{i=t-1}^{0}\Post_i \land
%%$~\qquad\qquad\,\displaystyle{\bigwedge_{i=0}^{m-1}(\beta_{1{n_i}{j_0}}=\alpha_i)\subst{\bigotimes_{j={n_i}-1}^{0}\Post\,'_{1j}}\land(\beta_{2{j_0}}=\alpha_m) \subst{\bigotimes_{j=m_1}^{0}\Post\,'_{1j}}\land}
%\end{split}
%\end{equation*}
%%$~\qquad\qquad\,\displaystyle{\bigwedge_{i=m+1}^{n}(\beta_{3{n'_i}{j_0}}=\alpha_i)\subst{\bigotimes_{j=n'_i-1}^{0}\Post\,'_{3j}\shortotimes\Post\,'_2\shortotimes\bigotimes_{j=m_1}^{0}\Post\,'_{1j}}\Big)}$\\
%\end{small}

\begin{footnotesize}
\begin{align*}
\Pred\,''&=\Big(\bigwedge_{i=0}^{m_1}\Pred\,'_{1i}\subst{\bigotimes_{j=i-1}^{0}\Post\,'_{1j}}\land\Pred\,'_2 \subst{\bigotimes_{j=m_1}^{0}\Post\,'_{1j}}\land\\ &\qquad
~\quad\bigwedge_{i=0}^{m_3}\Pred\,'_{3i}\subst{\bigotimes_{j=i-1}^{0}\Post\,'_{3j}\shortotimes\Post\,'_2\shortotimes\bigotimes_{j=m_1}^{0}\Post\,'_{1j}}\Big)\\
&=\Big(\bigwedge_{p=0}^{m_1}\!\Pred_{1p}\subst{\!\!\bigotimes_{j=p-1}^{0}\!\!\Post_{1j}}\land\Pred_2\subst{\!\!\bigotimes_{p=m_1}^{0}\!\!\Post_{1p}}\land \bigwedge_{p=0}^{m_2}\!\Pred_{3p}\subst{\!\!\bigotimes_{j=p-1}^{0}\!\!\Post_{3j}\otimes\Post_2 \otimes\!\!\bigotimes_{p=m_1}^{0}\!\!\Post_{1p}}\Big)\\&\qquad
\land \bigwedge_{t=0}^{n} \Pred_t\bigotimes_{i=t-1}^{0}\Post_i%\\&\qquad
\land \Big( \bigwedge_{i=0}^{m-1}(\beta_{1{n_i}{j_0}}=\alpha_i)\subst{\bigotimes_{j={n_i}-1}^{0}\Post\,'_{1j}}\land(\beta_{2{j_0}}=\alpha_m) \subst{\bigotimes_{j=m_1}^{0}\Post\,'_{1j}}\land\\ &\qquad
~\quad\bigwedge_{i=m+1}^{n}(\beta_{3{n'_i}{j_0}}=\alpha_i)\subst{\bigotimes_{j=n'_i-1}^{0}\Post\,'_{3j}\shortotimes\Post\,'_2\shortotimes\bigotimes_{j=m_1}^{0}\Post\,'_{1j}}\Big)\displaybreak[0]\\
&=\Big(\bigwedge_{p=0}^{m_1}\!\Pred_{1p}\subst{\!\!\bigotimes_{j=p-1}^{0}\!\!\Post_{1j}}\big)\land\Pred_2\subst{\!\!\bigotimes_{p=m_1}^{0}\!\!\Post_{1p}}\land \bigwedge_{p=0}^{m_2}\!\Pred_{3p}\subst{\!\!\bigotimes_{j=p-1}^{0}\!\!\Post_{3j}\otimes\Post_2 \otimes\!\!\bigotimes_{p=m_1}^{0}\!\!\Post_{1p}}\Big)\\&\qquad 
\land \bigwedge_{t=0}^{n} \Pred_t\bigotimes_{i=t-1}^{0}\Post_i
\land \Big(\bigwedge_{i=0}^{m-1}(\beta_{i}=\alpha_i)\subst{\bigotimes_{j={i-1}}^{0}\Post_{j}}\land(\beta_m=\alpha_m )\subst{\bigotimes_{j=m}^{0}\Post_{j}}\land\\ &\qquad
~\quad\bigwedge_{i=m+1}^{n}(\beta_i=\alpha_i)\subst{\bigotimes_{j=i-1}^{m}\Post_j\shortotimes\Post_m\shortotimes\bigotimes_{j=m-1}^{0}\Post_{j}}\Big)\\
&=\Pred
\end{align*}
\end{footnotesize}

>>>>

\Post\,'=\Post\uplus\bigotimes_{p=n}^{0}
		\Post_p
\end{mathpar}
Then, we have
	\[ P[Q]_{j_0}  
	\models
	{\openrule
		{
			\gamma_j^{j\in (J\setminus\{j_0\}) \uplus J_Q}, 
			\Pred\,',  \Post\,'
			 }
		{\ostate{s_i^{i\in L\uplus L_Q}} \OTWeakarrow {\alpha}
			\ostate{s_i'^{\, i\in L\uplus L_Q}}}
	}
	\]
\end{lemma}
~~\\
\begin{proof} Suppose we have:
\begin{mathpar}
%\[
P\models{\openrule
	{
		\gamma_j^{j\in J}, 
		\Pred,  
		\Post}
	{\ostate{s_i^{i\in L}} \OTWeakarrow {\alpha}
		\ostate{s_i'^{\, i\in L}}}
}
\end{mathpar}
By Lemma \ref{lem-rel-OT-WOT} this implies the following: \\			
%\begin{mathpar}
\begin{small}
$\forall p\!\in\![0\,..\,m_1].P\models\openrule
    {{\beta_{1pj}^{j\in J_{1p}}},\Pred_{1p},\Post_{1p}   }
         {\ostate{{(s_{1pi})}^{i\in L}} \OTarrow {\tau} \ostate{{(s_{1(p+1)i})}^{i\in L}}},  \quad
P\models\openrule
         {\beta_{2j}^{j\in J_2},\Pred_2,\Post_2 }
         {\ostate{({s_{20i}})^{i\in L}} \OTarrow {\alpha'} \ostate{({s_{21i}})^{i\in L}}} $ 
\\
and  $\forall p \in [0\,..\,m_2].P\models\openrule
         {
           \beta_{3pj}^{j\in J_{3p}},\Pred_{3p}, \Post_{3p}}
         { \ostate{(s_{3pi})^{i\in L}} \OTarrow {\tau} \ostate{(s_{3(p+1)i})^{i\in L}}}$
\end{small}
%\end{mathpar}

@@

However it must be noticed that in steps 1 and 3, we have two kinds of WOTs with different signatures (depending on whether $Q$ moves or not). It is still possible to glue them together in a global rule with two more terms for $\Pred$ and $\Post$ terms. This global merge is possible because the post-conditions of $P$ only act on variables of $P$ and those of $Q$ on variables of $Q$ (for example $\Post_i$ has no effect on $\Pred_{1p}$ and thus does not need to be taken into account when dealing with WOTs where $Q$ does not move).

We now  compare each element of the obtained WOT with the conclusion of the lemma:
{\small \begin{align*}
\alpha''&=\alpha'\displaystyle{\subst{\bigotimes_{j=m_1}^{0}\Post\,'_{1j}}}\\
& = \alpha'\displaystyle{\subst{\bigotimes_{j=m_1}^{0}\Post_{1j}}}&\text{$\alpha'$ only contains variables of $P$ untouched by $\Post_i$}\\
& = \alpha
\end{align*}}

For $\set{\gamma'}$ we distinguish elements in the holes of $P$ and of $Q$.

 First suppose $j\in J\setminus\{j_0\}$ we have $\gamma'_j=\gamma_j$ because $\Post'_{ij}$ has no effect on variables of $P$ and on $\beta_{1pj}$, consequently\\
{\small \begin{equation*}
\begin{split}\gamma'_j=&\mybigdotcup_{i=0}^{m_1} \vis{{\beta_{1ij}}\subst{\bigotimes_{k=i-1}^{0}\! \Post_{1k} } }  \dotcup  \vis{{\beta_{2j}}\subst{\bigotimes_{k=m_1}^{0}\!\Post_{1k}}}\, \dotcup
\mybigdotcup_{i=0}^{m_2} \vis{\beta_{3ij} \subst{\bigotimes_{k=i-1}^{0}\Post_{3k}\shortotimes\Post_2\shortotimes\bigotimes_{k=m_1}^{0}\Post_{1k}}}
\end{split}
\end{equation*}
}

Second, when $j\in J_t$ for some $t$,  $\gamma'_j$ is the concatenation of elements of $\gamma'_{1ij}$, $\gamma'_{2 j}$, $\gamma'_{3ij}$  that are not empty. By construction the concatenation of these elements is $\gamma_{tj}$, for $t\in[0..n]$. $\Post_{ik}$ has no effect on $\gamma_{tj}$ but  $\Post_{k}$ has.
%\TODO{do we want to be more explicit in the explanation? giving the exact indices is cumbersome because Post'ij is either empty or postij for the same indices as gamma, this can also be related with the ni and n'i}
We obtain:
{\small\begin{equation*}
\begin{split}
\gamma'_j =&
\mybigdotcup_{i=0}^{m_1}{\gamma'_{1ij}}\subst{\bigotimes_{k=i-1}^{0}\Post\,'_{1k}}\dotcup
 {\gamma'_{2 j}}\subst{\bigotimes_{k=m_1}^{0}\Post\,'_{1k}}\,\dotcup\, \mybigdotcup_{i=0}^{m_2}{\gamma'_{3ij}}\subst{\bigotimes_{k=i-1}^{0}\Post\,'_{3k}\shortotimes\Post\,'_2\shortotimes\bigotimes_{k=n}^{0}\Post\,'_{1k}}\\
=&
{\mybigdotcup_{t=0}^{n} {\gamma_{tj}}\subst{\bigotimes_{k=t-1}^{0} \Post_{k}}}   
\end{split}
\end{equation*}
}

Concerning predicates, we also separate predicates on $P$ from predicates on $Q$, and from the equality on the action filling the hole:
%\begin{small}
%\begin{equation*}
%\begin{split}
%\Pred\,''&=\bigwedge_{i=0}^{m_1}\Pred\,'_{1i}\subst{\bigotimes_{j=i-1}^{0}\Post\,'_{1j}}\land\Pred\,'_2 \subst{\bigotimes_{j=m_1}^{0}\Post\,'_{1j}}\land\\
%&~~{\bigwedge_{i=0}^{m_3}\Pred\,'_{3i}\subst{\bigotimes_{j=i-1}^{0}\Post\,'_{3j}\shortotimes\Post\,'_2\shortotimes\bigotimes_{j=m_1}^{0}\Post\,'_{1j}}}\\
%&=\big(\bigwedge_{p=0}^{m_1}\!\Pred_{1p}\subst{\bigotimes_{j=p-1}^{0}\Post_{1j}}\land\Pred_2\subst{\bigotimes_{p=m_1}^{0}\Post_{1p}}\land\\
%&~~~\bigwedge_{p=0}^{m_2}\!\Pred_{3p}\subst{\bigotimes_{j=p-1}^{0}\Post_{3j}\otimes\Post_2 \otimes\bigotimes_{p=m_1}^{0}\Post_{1p}}\big) \land \bigwedge_{t=0}^{n} \Pred_t\bigotimes_{i=t-1}^{0}\Post_i \land
%%$~\qquad\qquad\,\displaystyle{\bigwedge_{i=0}^{m-1}(\beta_{1{n_i}{j_0}}=\alpha_i)\subst{\bigotimes_{j={n_i}-1}^{0}\Post\,'_{1j}}\land(\beta_{2{j_0}}=\alpha_m) \subst{\bigotimes_{j=m_1}^{0}\Post\,'_{1j}}\land}
%\end{split}
%\end{equation*}
%%$~\qquad\qquad\,\displaystyle{\bigwedge_{i=m+1}^{n}(\beta_{3{n'_i}{j_0}}=\alpha_i)\subst{\bigotimes_{j=n'_i-1}^{0}\Post\,'_{3j}\shortotimes\Post\,'_2\shortotimes\bigotimes_{j=m_1}^{0}\Post\,'_{1j}}\Big)}$\\
%\end{small}

\begin{footnotesize}
\begin{align*}
\Pred\,''&=\Big(\bigwedge_{i=0}^{m_1}\Pred\,'_{1i}\subst{\bigotimes_{j=i-1}^{0}\Post\,'_{1j}}\land\Pred\,'_2 \subst{\bigotimes_{j=m_1}^{0}\Post\,'_{1j}}\land%\\ &\qquad
%~\quad
\bigwedge_{i=0}^{m_3}\Pred\,'_{3i}\subst{\bigotimes_{j=i-1}^{0}\!\Post\,'_{3j}\shortotimes\Post\,'_2\shortotimes\bigotimes_{j=m_1}^{0}\!\Post\,'_{1j}}\Big)\\
&=\Big(\bigwedge_{p=0}^{m_1}\!\Pred_{1p}\subst{\!\!\bigotimes_{j=p-1}^{0}\!\!\Post_{1j}}\land\Pred_2\subst{\!\!\bigotimes_{p=m_1}^{0}\!\!\Post_{1p}}\land \bigwedge_{p=0}^{m_2}\!\Pred_{3p}\subst{\!\!\bigotimes_{j=p-1}^{0}\!\!\Post_{3j}\otimes\Post_2 \otimes\!\!\bigotimes_{p=m_1}^{0}\!\!\Post_{1p}}\Big)\\&\qquad
\land \bigwedge_{t=0}^{n} \Pred_t\bigotimes_{i=t-1}^{0}\Post_i%\\&\qquad
\land \Big( \bigwedge_{i=0}^{m-1}(\beta_{1{n_i}{j_0}}=\alpha_i)\subst{\bigotimes_{j={n_i}-1}^{0}\Post\,'_{1j}}\land(\beta_{2{j_0}}=\alpha_m) \subst{\bigotimes_{j=m_1}^{0}\Post\,'_{1j}}\land\\ &\qquad
~\quad\bigwedge_{i=m+1}^{n}(\beta_{3{n'_i}{j_0}}=\alpha_i)\subst{\bigotimes_{j=n'_i-1}^{0}\Post\,'_{3j}\shortotimes\Post\,'_2\shortotimes\bigotimes_{j=m_1}^{0}\Post\,'_{1j}}\Big)\displaybreak[0]\\
&=\Big(\bigwedge_{p=0}^{m_1}\!\Pred_{1p}\subst{\!\!\bigotimes_{j=p-1}^{0}\!\!\Post_{1j}}\big)\land\Pred_2\subst{\!\!\bigotimes_{p=m_1}^{0}\!\!\Post_{1p}}\land \bigwedge_{p=0}^{m_2}\!\Pred_{3p}\subst{\!\!\bigotimes_{j=p-1}^{0}\!\!\Post_{3j}\otimes\Post_2 \otimes\!\!\bigotimes_{p=m_1}^{0}\!\!\Post_{1p}}\Big)\\&\qquad 
\land \bigwedge_{t=0}^{n} \Pred_t\bigotimes_{i=t-1}^{0}\Post_i
\land \Big(\bigwedge_{i=0}^{m-1}(\beta_{i}=\alpha_i)\subst{\bigotimes_{j={i-1}}^{0}\Post_{j}}\land(\beta_m=\alpha_m )\subst{\bigotimes_{j=m}^{0}\Post_{j}}\land\\ &\qquad
~\quad\bigwedge_{i=m+1}^{n}(\beta_i=\alpha_i)\subst{\bigotimes_{j=i-1}^{m}\Post_j\shortotimes\Post_m\shortotimes\bigotimes_{j=m-1}^{0}\Post_{j}}\Big)\\
&=\Pred
\end{align*}
\end{footnotesize}
